Master's Project Journal

### 2019-04-20 thesis/NetworkImplemenation.tex
79git statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit statusgit status

\subsubsection{\acs{ADPLL} 1 Design Summary}
The listed components were brought together in \textsc{NetworkADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_adpll}. Important to note is that each \ac{ADPLL} only contains two \acl{PD}s despite during network operation the need for phase comparisons to be made with up to four neighbours. The two ``missing'' \ac{PD}s are in fact synthesised as part of the other \ac{ADPLL}s, with which the comparison would be made. This allows for a 50\% reduction in the number of \ac{PD}s required, while also ensuring identical error is received by both \ac{ADPLL}s. Each \ac{ADPLL} performs comparison with the \ac{DCO} ``above'' and ``left'' of them in the Cartesian grid, and a port in the module outputs the results for use by the other \ac{ADPLL}s involved. %TODO juvenile wording
Table \ref{table:adpll1} contains a brief overview of the bit widths, tuning ranges and other configuration information for this \ac{ADPLL}.
\begin{table}[!h]
	\begin{center}
		\begin{tabular}{|l|r|r|r|}
			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
			\hline
			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
			\hline
			Frequency & $1.008~\si{\mega\hertz}$ & \multicolumn{1}{r|}{$62.988~\si{\kilo\hertz}$} & $8.944~\si{\mega\hertz}$ \T\\
			\hline
			Period & $992.0~\si{\nano\second}$ & \multicolumn{1}{r|}{$2.5~\si{\nano\second}$ (at $5~\si{\mega\hertz}$)} & $111.8~\si{\nano\second}$ \T\\
			\hline
		\end{tabular}
		\begin{tabular}{|l|r|l|r|}
			\multicolumn{4}{c}{Configuration as Implemented} \T\\
			\hline
			\ac{DCO} Counter Width & 12 bits & \ac{DCO} Control Width (max) & 12 bits \T\\
			\hline
			\ac{DCO} Bias Point & 79 & Runtime Gains & Enabled \T\\
			\hline
			Phase Detector Error Width & 7 & Error Sum Weight Width & 4 \T\\
			\hline
			\acs{kp} width & $2^{-5}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-8}\times[0,15]\cap\mathbb{Z}$ \T\\
			\acs{kp} & $2^{-5}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-8}\times[0,15]\cap\mathbb{Z}$ \T\\
			\hline
			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
			\hline
		\end{tabular}
	\end{center}
\caption[ADPLL Design 1 Summary]{ADPLL Design 1 Summary.}
\label{table:adpll1}
\end{table}

@@

Using this updated maximum number of inverters the frequency range can be recalculated. Firstly, the new minimum and maximum numbers of inverters are $309$ and $373$ respectively, with a mid point at $341$. The inverter delay and period step are then recomputed based on the centre frequency of $5~\si{\mega\hertz}$ as:
\begin{align}%TODO 2 x 2 x midpoint justification - in prev. chapter?
t_{step} &= \frac{T_{osc}}{0.5\times(n-\frac{k}{2})} = \frac{200\si{\nano\second}}{0.5\times 341} = 1.176~\si{\nano\second} \\
\tau_{inv} &= \frac{t_{step}}{2\times\text{inverters per step}} = \frac{t_{step}}{2\times2} = \frac{1.176~\si{\nano\second}}{4} = 293~\si{\pico\second}
\end{align}
The frequency range is then:
\begin{align} 
f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{(n-2CC)\times 2\times\tau_{inv}} \\
f_{min} &= \frac{1}{T_{max}} = \frac{1}{(373-0)\times 2\times 293~\si{\pico\second}} = 4.571~\si{\mega\hertz} \\
f_{max} &= \frac{1}{T_{min}} = \frac{1}{(373-2\times32)\times 2\times 293~\si{\pico\second}} = 5.518~\si{\mega\hertz}
\end{align}

The module implementing the \ac{RO} can be found in \textsc{RingOsc.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:ring_osc}.

@@

\section{Measurement Setup}\label{section:measurement_setup}
For the following sections in which the performance of either individual \ac{ADPLL}s or \ac{ADPLL} networks the same measurement setup was used. The \ac{FPGA} used was a Xilinx Artix-7 \acl{Nexys} on a Digilent \acs{Nexys} board. The \acl{Nexys} features 15850 logic slices, each with 4 lookup tables and 8 flip-flops. There are six clock management tiles, 240 dedicated digital signal processing slices and an analog-to-digital converter. The maximum operating frequency is $464~\si{\mega\hertz}$ and 210 input/output ports are available for use. Further information can be found in the datasheet \cite{a7_datasheet}. The \acs{Nexys} is an evaluation/development board for the aforementioned \ac{FPGA}, however directed at the education sector. It contains useful peripherals such as 16 switches, 8 7-segment displays and 4 user addressable \ac{PMOD} headers and can be used with the free of charge ``Webpack'' version of Vivado \cite{n4_datasheet}.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\textwidth]{../n4}
	\caption[\acs{Nexys} development board]{\acs{Nexys} development board.}
	\label{fig:n4}
\end{figure}

Measurements were performed using an Agilent MSO7054A, a four channel, $4~\si[per-mode=symbol]{\giga\sample\per\second}$ oscilloscope. Signals were extracted over the regular \ac{PMOD} headers as the board does not provide outputs better suited to the measurement of signals that are available to the user, such as SMA or BNC connectors, and the \ac{PMOD} header for the analog-to-digital converter, which has lower impedance traces, is not available for the designer to re-configure. Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si{\giga\sample\per\second}$. One of the four channels was consumed by the external reference, leaving three free for signal measurements, in most cases the generated output of three \ac{ADPLL}s were connected. $200\si{\micro\second}$ long captures were saved to an Agilent specified binary format which could be opened in Matlab for analysis.
%TODO diagram or picture of some sort

\section{\acs{ADPLL} Characterisation}
%TODO why each PLL

@@

\section{Minor Variations}\label{section:minor_variations}
%TODO osc size

\section{\acs{ADPLL} Network Implementation}
\subsection{2x2 Network}
\subsection{3x3 Network}

>>>>

\subsubsection{\acs{ADPLL} 1 Design Summary}
The listed components were brought together in \textsc{NetworkADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_adpll}. Important to note is that each \ac{ADPLL} only contains two \acl{PD}s despite during network operation the need for phase comparisons to be made with up to four neighbours. The two ``missing'' \ac{PD}s are in fact synthesised as part of the other \ac{ADPLL}s, with which the comparison would be made. This allows for a 50\% reduction in the number of \ac{PD}s required, while also ensuring identical error is received by both \ac{ADPLL}s. Each \ac{ADPLL} performs comparison with the \ac{DCO} ``above'' and ``left'' of them in the Cartesian grid, and a port in the module outputs the results for use by the other \ac{ADPLL}s involved. %TODO juvenile wording
Table \ref{table:adpll1} contains a brief overview of the bit widths, tuning ranges and other configuration information for this \ac{ADPLL}.
\begin{table}[!h]
    \begin{center}
        \begin{tabular}{|l|r|r|r|}
            \multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
            \hline
            \multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
            \hline
            Frequency & $1.008~\si{\mega\hertz}$ & \multicolumn{1}{r|}{$62.988~\si{\kilo\hertz}$} & $8.944~\si{\mega\hertz}$ \T\\
            \hline
            Period & $992.0~\si{\nano\second}$ & \multicolumn{1}{r|}{$2.5~\si{\nano\second}$ (at $5~\si{\mega\hertz}$)} & $111.8~\si{\nano\second}$ \T\\
            \hline
        \end{tabular}
        \begin{tabular}{|l|r|l|r|}
            \multicolumn{4}{c}{Configuration as Implemented} \T\\
            \hline
            \ac{DCO} Counter Width & 12 bits & \ac{DCO} Control Width (max) & 12 bits \T\\
            \hline
            \ac{DCO} Bias Point & 79 & Runtime Gains & Enabled \T\\
            \hline
            Phase Detector Error Width & 7 & Error Sum Weight Width & 4 \T\\
            \hline
            \acs{kp} width & 5 & \acs{ki} & 8 \T\\
            \hline
            \acs{kp} & $2^{-5}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-8}\times[0,15]\cap\mathbb{Z}$ \T\\
            \hline
            Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
            \hline
        \end{tabular}
    \end{center}
\caption[ADPLL Design 1 Summary]{ADPLL Design 1 Summary.}
\label{table:adpll1}
\end{table}

@@

Using this updated maximum number of inverters the frequency range can be recalculated. Firstly, the new minimum and maximum numbers of inverters are $309$ and $373$ respectively, with a mid point at $341$. The inverter delay and period step are then recomputed based on the centre frequency of $5~\si{\mega\hertz}$ as:
\begin{align}
t_{step} &= \frac{T_{osc}}{0.5\times(n-\frac{k}{2})} = \frac{200\si{\nano\second}}{0.5\times 341} = 1.176~\si{\nano\second} \\
\tau_{inv} &= \frac{t_{step}}{2\times\text{inverters per step}} = \frac{t_{step}}{2\times2} = \frac{1.176~\si{\nano\second}}{4} = 293~\si{\pico\second}
\end{align}
The frequency range is then:
\begin{align} 
f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{(n-2CC)\times 2\times\tau_{inv}} \\
f_{min} &= \frac{1}{T_{max}} = \frac{1}{(373-0)\times 2\times 293~\si{\pico\second}} = 4.571~\si{\mega\hertz} \\
f_{max} &= \frac{1}{T_{min}} = \frac{1}{(373-2\times32)\times 2\times 293~\si{\pico\second}} = 5.518~\si{\mega\hertz}
\end{align}

The module implementing the \ac{RO} can be found in \textsc{RingOsc.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:ring_osc}. In the module instance statement it can be seen that a second generated output is also present. This second output is taken from a point in the fixed part of the chain, a significant number of inverters prior to the minimum possible due to control code adjustment. This signal is at an identical frequency to that of the generated signal, however will not be immediately affected by a change in the control code, which could in some edge cases change the polarity of the signal. This signal is used to drive the \ac{LF} in order to avoid an extraneous addition to the accumulator in the integral path. This early signal must be taken from an inverter at an index significantly lower than $n-k$ to prevent this addition causing the problem it has been implemented to avoid.

@@

\section{Measurement Setup}\label{section:measurement_setup}
For the following sections in which the performance of either individual \ac{ADPLL}s or \ac{ADPLL} networks the same measurement setup was used. The \ac{FPGA} used was a Xilinx Artix-7 \acl*{Nexys} on a Digilent \acs*{Nexys} board. The \acl*{Nexys} features 15850 logic slices, each with 4 lookup tables and 8 flip-flops. There are six clock management tiles, 240 dedicated digital signal processing slices and an analog-to-digital converter. The maximum operating frequency is $464~\si{\mega\hertz}$ and 210 input/output ports are available for use. Further information can be found in the datasheet \cite{a7_datasheet}. The \acs*{Nexys} is an evaluation/development board for the aforementioned \ac{FPGA}, however directed at the education sector. It contains useful peripherals such as 16 switches, 8 7-segment displays and 4 user addressable \ac{PMOD} headers and can be used with the free of charge ``Webpack'' version of Vivado \cite{n4_datasheet}.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{../n4}
    \caption[\acs*{Nexys} development board]{\acs*{Nexys} development board.}
    \label{fig:n4}
\end{figure}

Measurements were performed using an Agilent MSO7054A, a four channel, $4~\si[per-mode=symbol]{\giga\sample\per\second}$ oscilloscope. Signals were extracted over the regular \ac{PMOD} headers as the board does not provide outputs better suited to the measurement of signals that are available to the user, such as SMA or BNC connectors, and the \ac{PMOD} header for the analog-to-digital converter, which has lower impedance traces, is not available for the designer to re-configure. Data was acquired with the oscilloscope in quad channel mode at a sampling rate of $2\si[per-mode=symbol]{\giga\sample\per\second}$. One of the four channels was consumed by the external reference, leaving three free for signal measurements, in most cases the generated output of three \ac{ADPLL}s were connected. $200\si{\micro\second}$ long captures were saved to an Agilent specified binary format which could be opened in Matlab for analysis.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\textwidth]{../measurement_setup}
	\caption[Measurement setup]{Measurement setup.}
	\label{fig:setup}
\end{figure}

\section{\acs{ADPLL} Characterisation}
Before combining the \ac{ADPLL}s in a network it was important to analyse the designs locked to an ideal reference, in order to establish a baseline performance. These tests were carried out with the \ac{ADPLL}s locked to an external reference at $5~\si{\mega\hertz}$. As \ac{ADPLL} 2 \& 3 have significant variance in the step sizes, for these oscillators range and step based measurements were obtained by measuring a number of oscillators, and typical values will be shown. Jitter was calculated based on the performance on nine individual \ac{ADPLL}s all locked simultaneously to the $5~\si{\mega\hertz}$ external reference. While jitter and skew are the key indicators of \ac{ADPLL} network performance, when it comes to the individual \ac{ADPLL}s themselves it is important to check lock and capture ranges also, to check that the implemented designs match with their intended values. The lock and capture ranges are given based on typical values, whereas jitter and skew are the median values from the \ac{ADPLL}s tested.

The \ref{table:characterisation} contains the results of this analysis. The first thing to note is that to the $\si{\kilo\hertz}$ the lock and capture ranges of the \ac{ADPLL}s implementing \ac{RO}s are identical. As the lock range would be expected to be slightly wider than the capture range in a design limited by the loop filter, it can be deduced that the limiting factor in the tuning range of the \ac{ADPLL} is the \ac{DCO} frequency range. This is confirmed by a cursory comparison against the values in Table \ref{table:adpll2}, where the \ac{RO} range is given as $4.571-5.518~\si{\mega\hertz}$. This range of $0.947~\si{\mega\hertz}$ is almost identical to those of the \ac{ADPLL} $1$ \& $3$ which confirms the \ac{RO} is the limiting factor. In \ac{ADPLL} 1 however, the lock and capture ranges show some difference, with the lock range being slightly wider. The lower end of both ranges is identical, as rather than the instant loss of lock seen at all other thresholds, the low end of both lock and capture ranges saw a more gradual degradation, and $1.08~\si{\mega\hertz}$ is the point that the lock was observed to be lost.

\begin{table}[]
    \begin{center}
        \begin{tabular}{lrrr}
        \multicolumn{1}{c}{}          & \multicolumn{1}{c}{ADPLL Design 1}  & \multicolumn{1}{c}{ADPLL Design 2} & \multicolumn{1}{c}{ADPLL Design 3} \\ \cline{2-4}       
        Capture Range                 & $1.08-8.61~\si{\mega\hertz}$       & $4.869-5.877~\si{\mega\hertz}$    & $4.377-5.363~\si{\mega\hertz}$       \T\\
        Lock Range                    & $1.08-8.635~\si{\mega\hertz}$      & As above                           & As above                            \\
        Jitter                        & $1.8385~\si{\nano\second}$         & $0.5623~\si{\nano\second}$         & $0.7410~\si{\nano\second}$          \\
        Skew (Mean \acs{TIE} to Ref.) & $1.7439~\si{\nano\second}$         & $0.3741~\si{\nano\second}$         & $0.6381~\si{\nano\second}$          \\
        Maximum \acs{TIE}             & $7.2840~\si{\nano\second}$         & $6.4770~\si{\nano\second}$         & $6.4578~\si{\nano\second}$          \\
        $k_p$                         & $\frac{1}{32}$                     & $\frac{1}{16}$                     & $\frac{1}{16}$                      \\
        $k_i$                         & $\frac{1}{256}$                    & $\frac{1}{64}$                     & $\frac{1}{64}$  
        \end{tabular}
    \end{center}
    \caption[\ac{ADPLL} Design Comparison]{\ac{ADPLL} Design Comparison.}
    \label{table:characterisation}
\end{table}

Looking then at the jitter and skew data, unsurprisingly the entirely \ac{FPGA} clocked \ac{ADPLL} with its larger period step is has the worst jitter, by a significant amount. The maximum \ac{TIE} is also correspondingly worse, which unless significant skew was present in the other \ac{ADPLL} designs is also to be expected. \ac{ADPLL} 2 \& 3 experience low jitter, at less than $1\%$ of the period. Somewhat surprisingly there is no apparent benefit in terms of jitter reduction due to the better delay resolution of the inverter based \ac{TDL}. This may be down to either the increase in jitter from a second source of variation, or a less favourable fractional-N synthesis due to the distribution of periods.

Despite Vivado being forced to use the same layout for both \ac{ADPLL} 2 \& 3 through the use of fixed cells, the minimum and maximum frequencies are still significantly different, although the size of the range is almost identical. This highlights the difficulty of ensuring identical conditions between measurements which is the most significant downside of inverter based modules.

@@

\section{\acs{ADPLL} Network Implementation}
As the \ac{ADPLL}s were designed with network usage in mind, the \ac{HDL} aspect of their integration into a network is straightforward. For \ac{ADPLL} 1, the oscillators need only be wired together, but for Design 2 \& 3 which are reliant on inverters require some care to ensure their lock ranges overlap sufficiently. This is a time consuming task, as a number of iterations may be required, even when locking down the cells of lined up \ac{RO}s, as the nets forming the connections may change. It is recommended only to carry out alignment when all other aspects of the design are unlikely to change and to save the bitstream generated. The projected time requirement in aligning \ac{RO}s meant that the creation of a 4x4 network was abandoned. \ac{FPGA} clocked \ac{ADPLL}s are more suited to larger networks as their centre point can be determined with ease.

The network operates using an external reference which is input over a \ac{PMOD} header that is shared with a number of output signals, however it could be configured to use an internal reference generated by either the \ac{FPGA} clock management utility or a \ac{DCO} of the designer's choice.

In order to increase the flexibility of the network, the switches on the \acs*{Nexys} are used to change both the proportional and integral gains and the configuration of the network at runtime. The reconfigurability aspect of the error combiner allows for the network to operate in either uni- or bi-directional mode, or as a number of individual \ac{PLL}s all connected to the external reference. This is particularly advantageous for the inverter based designs, as multiple combinations can be measured without reimplementing, and thus changing the layout of either oscillator or \acl{PD}.

Two different network sizes were implemented, a 2x2 and a 3x3 grid respectively. The 2x2 network was implemented first, and then this design extended to a 3x3 once measurements had been carried out on and any errors made had been corrected. In later sections matrix indexing will be used to refer to each oscillator, with \ac{ADPLL}$_{1,1}$ closest to the reference, and higher numbers in either index getting progressively further from the reference, as in Figure \ref{fig:network_mine}. A 3x3 network was the minimum goal, with the aforementioned but eventually not implemented 4x4 a stretch-goal if time permitted. A 3x3 network set as the primary goal, as it is the minimum size network in which there exists an \ac{ADPLL} with four neighbours. As will be explained in Chapter \ref{chap:5}, \ac{ADPLL} Design 1 was only implemented in a 2x2 configuration while both other \ac{ADPLL}s were tested in a 3x3 configuration.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{../network_annotated}
	\caption[ADPLL indexing explanation]{ADPLL indexing explanation.}
	\label{fig:network_mine}
\end{figure}

%TODO why use each PLL
